---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into the following options - under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus - and calculates how many loans are in each group. How would you describe the distribution of loans in a sentence?
Hint: You will need to use `case_when` to make this work.
**A1** The vast majority of loans are in the $10k to $50k range, only a few break into the 6 and 7 figure range.

```{r}
# case_when example
west_virginia_amount <- west_virginia_ppp %>%
mutate(
  amount = case_when(
    amount < 10000 ~ 'under_10k',
    amount >= 10000 & amount < 50000 ~ '10k_50k',
    amount >= 50000 & amount < 100000 ~ '50k_100k',
    amount >= 100000 & amount < 500000 ~ '100k_500k',
    amount >= 500000 & amount < 1000000 ~ '500k_1m',
    amount >= 1000000 ~ '1m_plus'
  )
)
west_virginia_amount %>%
  group_by(amount) %>%
     summarise(
    count_loans = n()
   
   
  )
```

**Q2** Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Davis Trust Company and Miners & Merchants Bank

```{r}
west_virginia_undisbursed<-west_virginia_ppp%>% 
  filter(undisbursed_amount>0)

west_virginia_undisbursed%>%
  group_by(servicing_lender_name)%>%
  summarise(total=n()
           )

```

**Q3** What are some potential reasons why a loan might not be paid out? What questions does that circumstance raise for you?
**A3** The loans could have been flagged for possible fraud. Seeing as almost all of the undisbursed loans are tied to one lender I would really want to look into their lending practices and who their primary clients are. It strikes me as odd that only one other lender has disbursed loans and they only have one. 

**Q4** Which month in 2021 had the most loans approved?
**A4** March

```{r}

west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))

```

**Q5** How many loans went to West Virginia applicants that listed the maximum number of jobs retained (500)? Which one had the earliest approved date?
**A5** 10, J. AUSTIN HEATING & COOLING, INC.

```{r}

west_virginia_max_retained<-west_virginia_ppp%>% 
  select(date_approved,jobs_retained,name)%>%
  filter(jobs_retained==500)

west_virginia_max_retained%>%
group_by(name)

           


```
