---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("84c34e9fd2690050a6b2968bd44aa4085801c08f")
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
acs5 <- load_variables(2019, "acs5", cache = TRUE)
View(acs5)


```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
#Write code here to read the RDS file

lab_08_data<-read_rds("data/lab_08.rds")


```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first in your list of ACS variables.
**A1**
Baltimore City
```{r}
low_med_age<-acs5
  median_age_total <- get_acs(geography = "county", variables = "B01002_001", year = 2019)

  md_low_md_age<-median_age_total%>%
  filter(str_detect(NAME,"Maryland"))

  
  



```

**Q2** Which Maryland counties have a median age of at least 50? Write code to isolate just those jurisdictions.
**A2**
Talbot and Worcester counties
```{r}

md_over_50<-md_low_md_age%>%
  filter(estimate>50)
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3**
Prince George's County, 25.6
```{r}
spanish_speakers_acs5<-acs5
  spanish_speaking_teens <- get_acs(geography = "county", variables = c(spanishagerange = "B16004_004"), state = "MD", year = 2019)
teens_total<-get_acs(geography = "county", variables = c(spanishspeaking =  "B16007_002"), state = "MD", year = 2019)

spanish_speaking_teens%>%
 inner_join(teens_total, by=c("GEOID", "NAME")) %>%
  rename(estimate_spanish_age_range=estimate.x, estimate_total_age_range=estimate.y) %>%
  mutate(percentage_population = 
           ((estimate_spanish_age_range/estimate_total_age_range)*100))%>%
           arrange(desc(percentage_population))




```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4**
Prince George's and Montgomery are some of the most populated counties in the state, they also have the highest percentages of Spanish speaking 5-17 by far. The next closest is only 10% and then 8%. We see PG and MoCo leading by wide margins. Garret county only has 29 Spanish speakers 5-17, which is remarkably low. 


**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5**
60620, a zip code in Chicago which is about 96% black. 

```{r}
total_pop_by_zip <- get_acs(geography = "zcta", variables = c(pop = "B02001_001"), year = 2019)
black_pop_by_zip<-get_acs(geography = "zcta", variables = c(blackpop =  "B02001_003"), year = 2019)

percentage_black_pop<-total_pop_by_zip%>%
  inner_join(black_pop_by_zip, by=c("GEOID","NAME"))%>%
  rename(estimate_total_pop_by_zip=estimate.x, estimate_black_pop_by_zip=estimate.y)%>%
  mutate(percentage_pop = ((estimate_black_pop_by_zip/estimate_total_pop_by_zip)*100))%>%
  arrange(desc(percentage_pop))
  
ppp_loan_applications <- lab_08_data%>%
  mutate(zip5 = str_sub(zip, 1, 5))

top_zips<-ppp_loan_applications%>%
  group_by(zip5)%>%
  summarise(total=n())%>%
  arrange(desc(total))

removed_by_zip<-top_zips%>%
  left_join(percentage_black_pop, by=c("zip5"="GEOID"))%>%
  arrange(desc(total))

```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed applications in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6**

All but the second result are located on either the South or West Side of Chicago. The zip with the second most removed loans is in Atlanta, GA. Almost all of the zips have a population that is well over 50% black. A notable exception is zip 60623 on the West Side of Chicago which is only about 30% black. 

**Q7** What other columns from the PPP data could be useful in reporting on what happened to these removed loans?
**A7**
I would also like to pull the data for hispanic applicants to see how they compare. It would be helpful to see the information for when these loan applications were made too. Was this early in the pandemic or later? If there were data on median housefold income and percent with a high school diploma for the zips that would be good too. 
