---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [STORY NAME](STORY URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

fatal_police_shootings <-read_csv("fatal-police-shootings-data.csv")

fatal_police_shootings_clean <- read_csv("fatal-police-shootings-data-clean.csv")

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "Of men killed, 57 percent were armed with a gun and of women, 44 percent. In other cases, women were armed with knives, cars, toy weapons, hammers and hatchets."
* **Analysis summary**: 
We used open refine to clean up the gun column and get all of the incidents with guns. We then created separate data frames for each gender with incidents involving a gun. Then we calculated the percentages. Our results were not identical to the Post's and this may have had to do with them calculating it at a different time, but our results are close.
```{r}




all_police_shootings_by_gender <- fatal_police_shootings_clean%>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))

all_shootings_men <- all_police_shootings_by_gender %>% 
  filter(gender == 'M')

all_shootings_women <- all_police_shootings_by_gender %>% 
  filter(gender == 'F')

gun_police_shootings_by_gender <- fatal_police_shootings_clean %>%
  filter(armed_clean == "gun") %>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
  

all_shootings_men_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'M')

all_shootings_women_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'F')

men_with_gun_percentage <- ((4012/6814)*100)
women_with_gun_percentage <- ((145/320)*100)

men_with_gun_percentage
women_with_gun_percentage

```

### Sentence 2

* **Sentence text**: “Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. While about twice as many White women were shot and killed as Black women, White women account for five times the Black female population.” 
* **Analysis summary**: We've managed to replicate the Post's findings almost. However, we have different totals for white, Hispanic, and other races than they did. We get closer if we filter by age, but that inst still exact. We don't know what age they were defining as making someone a woman.
```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

shootings_women_race<-fatal_police_shootings%>%
  filter(armed == "unarmed", gender=="F", age >17)%>%
  group_by(armed, gender, race)%>%
  summarise(
    count=n()
    )

```

### Sentence 3

* **Sentence text**: "Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. "
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]


We removed all of the N/A results from race. Then we filtered by race and gender and sumarrized to get the total of black women shot by police which is 54. Then we calculated the total shot which is 5941. We got 0.9% which replicates the Post's <1%.It is possible our result is higher if it has been changed since the initial result, but it's still in the given range.
```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock


shootings_race_known<-fatal_police_shootings%>%
 drop_na(race)

  shootings_black_women<-shootings_race_known%>%
  filter(gender == 'F', race == 'B' )%>%
  summarise(
    count=n()
    )
  
shootings_race_known%>%
   summarise(
    count=n()
    )

percent_black_women<- 
  ((54/5941)*100)

```

### Sentence 4

* **Sentence text**: Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. But within this small subset, Black women ... account for 20 percent of the women shot and killed and 28 percent of the unarmed deaths.”
* **Analysis summary**: We were able to recreate the first number assuming cases where race is unknown are excluded. We don't know whether those cases were excluded or not. There is a 2% difference in our finding on the percentage of unarmed deaths black women accounted for, but this is a reasonable margin of error especially if the dataset was updated after the story was published. The real confusing part is whether they meant 28% of unarmed women or 28% of all unarmed deaths. It only makes sense for it to be unarmed women, there are just too few deaths relative to the total, but the sentence used in the article makes it seem like they're saying unarmed deaths as a whole. 
```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

shootings_race_known<-fatal_police_shootings%>%
 drop_na(race)

  shootings_women<-shootings_race_known%>%
  filter(gender == 'F' )%>%
  summarise(
    count=n()
    )
  
percent_black_women_women<-
((54/280)*100)


shootings_unarmed_women<-fatal_police_shootings%>%
  filter(armed == "unarmed", gender=="F")%>%
  group_by(armed, gender, race)%>%
  summarise(
    count=n()
    )

percent_black_women_unarmed<-
((7/27)*100)


```

### Sentence 5

* **Sentence text**: Black men... make up 27 percent of the men shot and 36 percent of the unarmed deaths.
* **Analysis summary**: Again we exactly recreated the first result, but again our second result is 2 points off. It is really strange that it is exactly 2% in both cases. This is not from excluding race unknown, so why is it exactly 2% off in calculations for both women and men that are unarmed? The dataset could have changed, but the consistency seems odd. The vague wording in both cases could be to blame.  

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

shootings_race_known<-fatal_police_shootings%>%
 drop_na(race)

  shootings_men<-shootings_race_known%>%
  filter(gender == 'M')%>%
  summarise(
    count=n()
    )
  
    shootings_black_men<-shootings_race_known%>%
  filter(gender == 'M', race == 'B' )%>%
  summarise(
    count=n()
    )
 percent_black_men<- 
((1538/5660)*100)
 
 
 shootings_unarmed_men<-fatal_police_shootings%>%
  filter(armed == "unarmed", gender=="M")%>%
  group_by(armed, gender, race)%>%
  summarise(
    count=n()
    )
percent_unarmed_black_men<-
((133/389)*100)
 

```

-30-