---
title: "lab_11"
author: "sean mussenden"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

install.packages("vctrs", repos = "https://packagemanager.rstudio.com/cran/latest")
```
```{r}
install.packages(tidycensus)

```
## You will need

* A Census API key

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. You'll need tidyverse, tidycensus, janitor, sf and tigris, and you'll also need to install your census API key.

```{r}
options(scipen=999)
library(tidyverse)
library(tidycensus)
library(sf)
library(janitor)
library(tigris)
library(lubridate)
census_api_key("84c34e9fd2690050a6b2968bd44aa4085801c08f")
```

## Load data

**Task** Create a codeblock and load the WV slice of the PPP loan data, which is in the data folder and is called lab_11.rds.

```{r}
WV_slice<-read_rds("data/lab_11.rds")


```

## Questions

**Q1.** You are interested in doing some reporting on how West Virginia companies in the "Mining, Quarrying, and Oil and Gas Extraction" industry (https://www.naics.com/six-digit-naics/?code=21) used PPP loans during the pandemic (West Virginia is a big coal mining state).  

Then produce a choropleth map that shows West Virginia counties shaded according to the total amount of loans to that industry per 100,000 population.

Use the magma color palette, and shade on a "log" scale. Also produce a table that shows the total amount of loans per 100,000 population for each county, ranked from highest to lowest, along with the total amount of loans and the total number of loans.

Then answer these questions:

* Based on the map, what region of the state appears to have gotten more loans in this industry? What geographic pattern do you notice in the data?
* Based on the table, what county got the highest total loan amount? What was the total amount per 100K population, the total amount of loans (not population adjusted), how many loans were there, and what's the population?

```{r}
mining_quarrying <- WV_slice %>% 
   mutate(mining = str_sub(naics_code, 1, 2)) %>% 
  filter(mining == "21")

```

```{r}
  county_loans<-mining_quarrying%>%
  group_by(project_county_name)%>%
  summarise(total_loans=n(), loan_amount = sum(amount))%>% 
  arrange(desc(total_loans))
```

```{r}
county_pop<-get_acs(geography = "county",
  variables = c(population = "B01001_001"),
  state = "WV",
  geometry = TRUE)
```

```{r}
county_poppier <- county_pop %>%
  rename(COUNTY = NAME) %>%
  mutate(COUNTY = toupper(str_remove_all(COUNTY,", West Virginia|County"))) %>%
  mutate(COUNTY = str_trim(COUNTY,side="both")) %>%
  rename(population = estimate) %>%
  select(COUNTY, population, geometry)
```

```{r}
wv_loans_per100k <- county_poppier %>% 
  left_join(county_loans, by=c("COUNTY" = "project_county_name"))
```

```{r}
loans_per_100k<-wv_loans_per100k%>%
mutate(total_loans = replace_na(total_loans,0)) %>%
  mutate(loans_per_100k = loan_amount/population*100000) %>%
  arrange(
      desc(loans_per_100k))

```

```{r}
ggplot() + 
  geom_sf(data=loans_per_100k, aes(fill=loans_per_100k)) + 
  theme_minimal()+
  scale_fill_viridis_b(option="magma",trans = "log") 


```


**A1.**
Even though other regions actually received many more loans by the numbers, Ritchie County received far and away the most money per 100k. Even without the qualifier of per 100k people, Ritchie County Received a staggering sum of money. Ritchie County only has a population of 9,747, but it received $7,653,980.00 in PPP loans, which works out to $78,526,520.98, more than twice as much as second place Gilmer County. 

The fewest loans seem to have been sent to the Southwest and far Eastern corners of the state. It is lilely that relatively little mining goes on in these areas. 




**Q2.** You want to understand WHY the spatial pattern you saw above exists.  So you grab a shapefile with the latitude and longitude coordinates of every mine permit in West Virginia from the state Department of Environmental Protection.

Here's the url: https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json

(It's in a slightly different format than the .shp file in the lab. It's another type of shapefile called "geojson".  It will work just the same, though, despite that difference. You can read it directly into R without manually downloading it, by putting the url in the st_read() function we learned about in this week's prelab.)

This data includes both currently operating mines and old ones that have closed.  You want to examine only mines that have a permit expiration date after Sept. 1, 2021.

Make a map with a point for each one of these mine permits and put a blank map with the outline of West Virginia counties (which you can get from Tigris) behind it.

What spatial pattern do you see? How does it compare to the spatial pattern in the first map you made?


```{r}
i_me_mining<- st_read("https://tagis.dep.wv.gov/arcgis/rest/services/WVDEP_enterprise/mining_reclamation/MapServer/0/query?where=1%3D1&outFields=*&geometry=&geometryType=esriGeometryEnvelope&inSR=4326&spatialRel=esriSpatialRelIntersects&outSR=4326&f=json")
```

```{r}
i_me_mining <- i_me_mining %>% 
  mutate(expiredate=ymd(expiredate)) %>% 
  filter(expiredate > "2021-09-01")

wv_counties <- counties() %>% 
 filter(STATEFP == "54")

```
 
```{r}
 ggplot() + 
  geom_sf(data=wv_counties) + 
  geom_sf(data=i_me_mining) +
  theme_minimal()
```

**A2.**

So this paints a totally different picture than what I read from the first one. The majority of mines are in the Southwest of the state, running in a slightly curved band from Southwest to North Northeast. Now I'm very confused about what was going on in Ritchie County. 

**Q3.** To confirm the spatial comparison you identified in question 2, let's create a new map that layers the mine location points on top of our PPP choropleth map we created in question 1. What do you notice about the number of mines in the county that has the highest total loan amount per 100K?

```{r}
ggplot() + 
  geom_sf(data=wv_counties) + 
  geom_sf(data=loans_per_100k, aes(fill=loans_per_100k))+ 
  geom_sf(data=i_me_mining) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log") 

```


**A3.**

There is at most 1 mine in Ritchie County and it may be on in a another county, it's hard to tell. 

**Q4.** Starting from the original WV ppp dataframe, examine the loans to companies in "Mining, Quarrying, and Oil and Gas Extraction" in the county with the highest total loan amount per 100K (you can use the table you created in question 1 to identify the top county).

* What single company accounts for the biggest chunk of the total loan amount?
* What specific industry are they in, based on their 6-digit NAICS code? Look them up on the web. What services do they provide?
* How many jobs did they save, according to the data?
* What's your hypothesis for why a company in their specific industry would need to seek a PPP loan during the pandemic?

```{r}

Who_is_ritchie<- WV_slice %>% 
  mutate(ritchie_mining = str_sub(naics_code, start=1L, end=2L)) %>%
  filter(project_county_name == "RITCHIE") %>%
  filter(ritchie_mining =="21") %>%
  arrange(desc(amount))

Who_is_ritchie

```
**A4.**
Hall Drilling received the largest chunk of loan money. Their NAICS code, 213111, indicates that they are in the business of Drilling Oil and Gas Wells. According to their website, Hall Drilling is "...one of the most respected oil & gas companies in the Appalachian Basin with over 300 skilled industry employees."

The PPP data indicates Hall retained 387 jobs as a result of the loans received. This is in line with their claim to have "over 300 skilled industry employees." 

The early days of the pandemic were a terrible time for oil and gas. Almost no one was traveling. It can be hard to remember given the current petroleum climate, but at one point in 2020 oil was trading for -$37 a barrel. Oil producers literally had to pay people to take the oil away. A company like Hall Drilling probably was in dire straits in those days. 


-30-
