---
title: "Montana_profile"
author: "Hunter Savery"
date: '2022-05-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this assignment.

```{r}
install.packages("ggthemes")
library(tidycensus)
library(tidyverse)
library(tigris)
library(lubridate)
library(janitor)
library(ggthemes)
options(scipen=999)
```



**Q1.**. Write R code that generates some basic descriptive statistics that allows you to describe the applications from your state. This code should produce results that describe the data in different ways: how many applications there are, along with ways to understand the typical and most frequent values for columns you find interesting or newsworthy. You should produce at least five (5) descriptive statistics and write up a summary of the data using them.

```{r}
Montana <- read_csv("montana_ppp.csv")
```


```{r}
Montana <- Montana %>% 
  mutate(
    city = str_to_upper(city))
write_csv(Montana,"montana2.csv")
```

```{r}
Montana_clean <- read_csv("montana2.csv") %>% 
  mutate(zip = str_sub(zip, start=1L, end=5L))
```

```{r}
#number of PPP applications##
nrow(Montana)
```

```{r}
apps_by_city <- Montana_clean %>% 
  group_by(city) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))
```

```{r}
## average loan amount ##
mean(Montana_clean$amount)

```

```{r}
## % of loans that are forgiven ##
forgive_and_forget <- Montana_clean %>% 
  filter(forgiveness_amount >= amount)
(nrow(forgive_and_forget)/nrow(Montana_clean)) * 100
```

```{r}
## % of loans that are unforgiven ##
some_things_can_never_be_forgiven <- Montana_clean %>% 
  filter(forgiveness_amount < amount)
(nrow(some_things_can_never_be_forgiven)/nrow(Montana_clean)) * 100
```

```{r}
## % undisbursed loans##

undisbursed <- Montana_clean %>% 
  filter(undisbursed_amount > 0)
(nrow(undisbursed)/nrow(Montana_clean)) * 100
```


```{r}
## average number of jobs retained##

mean(Montana_clean$jobs_retained)

```



**A1.** 

Montana had 47,066 PPP loan applications. Urbanish centers Billings, Missoula, Bozeman, Helena, and Great Falls had the most applications. The average loan approval amount was $55,073.78. Of loans approved, 90.03% were forgiven in full. Just 6.01% were not forgiven. Impressively, 100% of loans have been disbursed. The average number of jobs retained was 6.89, whcih you can round up or down depending on your optimism level. 

## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. 

```{r}
load_variables(2019, "acs1")
Montana_pop <- get_acs(geography = "state", 
                  variables = "B01001_001", 
                  year = 2019, geometry = TRUE) %>% 
                  filter(NAME == "Montana")

county_pop <- get_acs(geography = "county",
                      variables = "B01001_001", 
                      year = 2019, geometry = TRUE) %>% 
                      filter(str_detect(NAME, ", Montana"))

zip_pop <- get_acs(geography = "zcta", 
                   variables = "B01001_001",
                   year = 2019, 
                   geometry = TRUE) %>% 
                  filter(GEOID >= 59001 & GEOID <= 59937)
```


```{r}


per_capita_loans <- Montana %>% 
  summarise(count=n()) %>% 
  mutate(per_capita = count/Montana_pop$estimate)



```


```{r}
county_poppier <- county_pop %>% 
  mutate(
    NAME = str_replace(NAME," County, Montana",""),
    NAME = str_to_upper(NAME)
  ) %>% 
  rename(
    project_county_name = NAME,
    population = estimate
  ) %>% 
  select(project_county_name,population,geometry)


zip_poppier <- zip_pop %>% 
  rename(
    population = estimate,
    zip = GEOID
  ) %>% 
  select(zip,population,geometry)
```



```{r}
county_ppp <- Montana_clean %>% 
  group_by(project_county_name) %>% 
  summarise(count=n()) %>% 
  left_join(county_poppier, by = "project_county_name")


zip_ppp <- Montana_clean %>% 
  group_by(zip) %>% 
  summarise(count=n()) %>% 
  left_join(zip_poppier, by = "zip")
```


```{r}
county_per_capita <- county_ppp %>% 
  mutate(loans_per_capita = count/population) %>% 
  arrange(desc(loans_per_capita))


zip_per_capita <- zip_ppp %>% 
  mutate(
    loans_per_capita = count/population,
    per_cap_diff = mt_loans_per_cap$per_capita - loans_per_capita
  ) %>% 
  arrange(desc(per_cap_diff))

Montana_per_capita <- Montana_pop %>% 
  select(NAME,estimate) %>% 
  mutate(
    loans = nrow(Montana_clean),
    loans_per_cap = loans/estimate
  )
```

Then, make a county map using ggplot showing the per-capita data and a zip code map showing the difference from the statewide per-capita figure.


```{r}
counties <- counties()
Montana_counties <- counties %>% 
  filter(STATEFP == "30")
Montana_counties %>% 
  ggplot() +
  geom_sf() +
  theme_minimal()
```


```{r}
ggplot() +
  geom_sf(data=Montana_counties, fill="white") +
  geom_sf(data=cou_per_cap, 
  aes(fill=loans_per_capita,geometry=geometry)) +
  scale_fill_gradient(low="yellow",high="red") +
  theme_minimal()
```

```{r}
ggplot() +
  geom_sf(data=mt_counties, fill="white") +
  geom_sf(data=zip_per_capita, 
          aes(fill=per_cap_diff,geometry=geometry)) +
  scale_fill_distiller(type = "div", palette = 1, direction = 1, aesthetics = "fill") +
  theme_minimal() +
  labs(
    title = "Areas with low populations received the most loans per capita, but that isn't saying much."
  )
```


Describe the most interesting or newsworthy findings based on your exploration.

**A2.** 
The first heat map, based on counties, shows more dramatic results for loans per capita. When we look at it by zip, we see that there isn't much of a disparity between different communities. Small, rural locales have the highest per capita rates of PPP loans, as opposed to the "sprawling metropolises" of Bozeman or Helena. It's interesting to see it this uniform, but then again not many people live in Montana. This only further entrenches my belief that Montana is the finest state in the union, political leadership aside. 

I'm now curious to see the disparities in other Western states. Does this pattern hold true in Wyoming or Idaho? Colorado would be an interesting comparison because it has more major urban centers compared to Montana. There may be more of an urban-rural divide. 


**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. Describe the most noticeable outlier lenders on that graphic below. 

#All loans disbursed. Way to go Montana! ##

If there are no undisbursed loans, make graphic showing the relationship between lenders and loan amounts showing how many loans each lender issued for each amount in the dataframe. Describe the most noticeable outlier lenders on that graphic.


```{r}
Take_me_to_your_lender <- Montana_clean %>% 
  mutate(
    loan_sizes = case_when(
      amount < 50000 ~ "sub50k",
      amount >= 50000 & amount < 100000 ~ "50k_to_100k",
      amount >= 100000 ~ "100kplus")
  ) %>% 
  group_by(servicing_lender_name,loan_sizes) %>% 
  summarise(
    count=n()
  ) %>%
  arrange(servicing_lender_name)
```


```{r}
Take_me_to_your_lender %>% 
  ggplot(aes(x=loan_sizes,y=count,label=servicing_lender_name)) +
  geom_point() +
  geom_text(aes(label=ifelse(count>1000,as.character(servicing_lender_name),
        '')),hjust=-0.05,vjust=0)
```


```{r}
offshore_bank_accounts <- Montana_clean %>% 
  filter(servicing_lender_state != "MT") %>% 
    mutate(
    loan_sizes = case_when(
      amount < 50000 ~ "sub50k",
      amount >= 50000 & amount < 100000 ~ "50k_to_100k",
      amount >= 100000 ~ "100kplus")
  ) %>% 

  group_by(servicing_lender_name,loan_sizes) %>% 
  summarise(
    count=n()
  ) %>%
  arrange(servicing_lender_name)
```

```{r}
offshore_bank_accounts %>% 
  ggplot(aes(x=loan_sizes,y=count,label=servicing_lender_name)) +
  geom_point() +
  geom_text(aes(label=ifelse(count>100,as.character(servicing_lender_name),
            '')),hjust=-0.05,vjust=0)
```


**A3.** 

So my hypothesis here did not hold water. I figured out-of-state lenders would be financing primarily high value loans. This didn't seem to be the case, the majority of loans were small regardless of whether the lender was based in or out of state. Among all lenders Glacier Bank is dominant across all categories, especially small loans. US Bank is the king of out-of-state lenders and its pattern maps onto Glacier's just in miniature. 


**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. 

To do this, you will need to add FIPS codes to your PPP dataset.


```{r}
naics_survey <- read_csv("cbp19co.csv") %>% 
  filter(
    fipstate == 30 &
    !str_detect(naics, "-") &
    !str_detect(naics, "/")
  ) %>% 
  mutate(
    fips5 = str_c(fipstate,fipscty),
    naics = as.numeric(naics)     
  ) %>% 
  select(fips5,naics,est)


naics_key <- read_csv("naics_codes.csv") %>% 
  rename(naics = naics_code)

naics_survey_names <- naics_survey %>% 
  left_join(naics_key, by = "naics")
```



```{r}
all_fips <- fips_codes %>% as_tibble()
mt_fips <- all_fips %>% filter(state == 'MT') %>% 
  mutate(
    county = str_replace(county, ' County', ''),
    project_county_name = str_to_upper(county),
    fips5 = str_c(state_code,county_code)
  ) %>% 
  select(project_county_name,fips5)
Montana_fips <- Montana_clean %>% 
  left_join(mt_fips, by = "project_county_name") %>% 
  rename(naics = naics_code)
```



```{r}
mt_naics <- Montana_fips %>% 
  group_by(naics) %>% 
  summarise(
    count=n()
  ) %>% 
  arrange(desc(count)) %>% 
  head(10)

statewide_cbp <- naics_survey_names %>% 
  group_by(naics,title) %>% 
  summarise(total = sum(est)) %>% 
  arrange(desc(total))


Montana_v_cbp <- mt_naics %>% 
  left_join(statewide_cbp, by = 'naics')%>% 
  mutate(
    diff = total - count
  )
```



```{r}
mt_naics_big_boys_only <- Montana_fips %>% 
  filter(
    business_type != "Sole Proprietorship" &
      business_type != "Self-Employed Individuals"
  ) %>% 
  group_by(naics) %>% 
  summarise(
    count=n()
  ) %>% 
  arrange(desc(count)) %>% 
  head(10)


Montana_x_cbp_big_boys_only <- mt_naics_big_boys_only %>% 
  left_join(statewide_cbp, by = 'naics') %>% 
  mutate(
    diff = total - count
  )
```


```{r}
gimme_a_drink <- Montana_fips %>% 
  filter(
    naics == 722410 &
      business_type != "Sole Proprietorship" &
      business_type != "Self-Employed Individuals"
  ) %>% 
  group_by(project_county_name,fips5,naics) %>% 
  summarise(count=n()) %>% 
  arrange(desc(count))


naics_survey_bars <- naics_survey %>% 
  filter(naics == 722410)

mt_bars_vs_cbp <- gimme_a_drink %>% 
  left_join(naics_survey_bars, by = "fips5") %>% 
  mutate(
    est = case_when(
      is.na(est) ~ 0,
      TRUE ~ est),
    diff = count - est)


write_csv(mt_bars_vs_cbp, "mt_bars_vs_cbp.csv")
```

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

**A4.** 

There are some significant gaps between the official count and actual number of loans disbursed across the top ten industries. However, two out of the top ten have no data, which is a big problem! 

I chose to look at bars, or as the census calls it drinking places (alcoholic beverages), because when I was in Montana last summer I sure went to a lot of bars. The census under counted a lot of bars in popular tourist areas, but the curious thing is Cascade County is an outlier. The Census Bureau missed a whopping 86 bars there. Cascade is not a hub of tourist activity and it doesn't seem reasonable that they are flooded with bars, so it's worth digging into what's going on there. 

https://datawrapper.dwcdn.net/c35Wk/1/

## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.**
Well, if I was writing for a Montana outlet I would want to know what the heck is going on in Cascade County. Are there an exceptional number of bars there? Did the census just really mess up this one area? Why? This is probably a relatively small story, but it's a thread worth pulling. 

I would want to find out how many bars are in Cascade County, the county demographics, what are the largest employers in the area. Is it tourism? Ranching? If you told me there were only 86 bars there I would believe you, so did the Census Bureau miss all of them? I would want to look at trends in other states. Where are bars undercounted? In general, it is high tourism areas, but are there other outliers like Cascade County in other states? 

At a minimum something funky happened in Cascade County, at a maximum the Census Bureau is very bad at counting bars.