---
title: "lab_01"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
install.packages('tidyverse')
```

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data.

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.
**Q1** How many rows and columns does this dataframe have?
**A1** 33, 336 rows, 58 columns

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

Use the pre-lab and your reference notebook as examples.

**Q2** Which West Virginia county got the most loans? How many?
**A2** Kanawaha, 4,273

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** Palm Beach, 2 (I love the concept of a Palm Beach County WV)

```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange((count_loans))
```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** Kanawha County is the most populous county in WV and the home of Charleston, the state capital and largest city, all of these make it a likely candidate to reeive the majority of PPP loans. As far as I can tell, Palm Beach County, WV doesn't exist. I'm not sure what's happening with these loans, probably an error. The next lowest is Allegheny County which is also not in WV, there may be issues with this data set. 


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** Greenbrier Hotel Corporation, $8,912,202.00, hotel, 500 jobs retained

```{r}
wv_ppp_loans %>%
  arrange(desc(amount))
```

**Q6** What West Virginia politician (with his family) owns the company identified in question 5? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** According to ProPublica, The Greenbrier is owned by Jim Justice and his family. Justice is both the state's governor and richest man.

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** The size of that loan to the Greenbrier is staggering. It also came to  county where the median loan amount was only $19,750.00, meaning this loan was orders of magnitude larger than what anyone else in the county was getting. Comparing that median loan size to Kanawha, which had the highest loan count, we see that it's lower, suggesting a significant wealth gap in Greenbrier County. 
```{r}
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
saveRDS(data,file="wv_summary.rds")
```

